\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[backend=biber,style=numeric-comp, sorting=none]{biblatex}
\usepackage{hyperref}
\usepackage{placeins}
\lstset{
  basicstyle=\ttfamily,
  columns=fullflexible,
}
\definecolor{light-gray}{gray}{0.95}
\addbibresource{ref.bib}
\title{ARGES: \\ Advanced Renormalization Group Equation Simplifier \\ User Manual}
\date{2017}
\author{Tom Steudtner}
\begin{document}
\begin{titlepage}
\maketitle
\end{titlepage}
\newpage
\tableofcontents
\thispagestyle{empty}
\newpage
\section{Motivation - Why ARGES?}
Arges is the name of a cyclops in the greek mythology, which was exiled by the Titans but freed by Zeus to forge lightning bolts for his battle against his former masters. \newline
In this tradition as a useful tool, ARGES - Advanced Renormalization Group Equation Simplifier is a framework to calculate perturbative renormalization group equations for renormalizable quantum field theories, using \cite{LuoWangXiao,MV1,MV2,MV3,vev1,vev2,gauge3L,gauge3L2,OsbornJack,OsbornJack2,OsbornJack3,Decoding}. Of course, other tools exist which offer an intersection of the same core features, most famously SARAH \cite{SARAH,SARAH4} and PyR@TE \cite{pyrate,pyrate2}. However, this piece of software tries to adress some of their limitations and adopts their resolution as its guiding features:
\begin{itemize}
\item it is possible to define particle numbers, representations and gauge group ranks as variables, even at all at once
\item gauge groups may be defined as totally unknown altogether
\item gauge invariants are not automatically resolved
\item for interaction terms, contractions of generation and gauge indices are completely configurable
\item disentanglement of parameter RGEs is up to the user, hence the program cannot fail while trying to resolve these
\end{itemize}
ARGES is limited to irreducible representation and RGEs in the $\overline{\mathrm{MS}}$ scheme, however, for all mass independent schemes, beta functions are universal at one loop (two loop for gauge couplings).
\newpage
\section{Prerequisites}
ARGES calculates $\overline{\mathrm{MS}}$ scheme Beta functions of the form:
\begin{align}
\beta_x = \frac{\partial x}{\partial \mathrm{ln}(\mu)}
\end{align}
The software assumes the following perturbative renormalizable Lagrangian to as a basis of its calculations:
\begin{align}
\mathcal{L} = &-\frac{1}{4} F^A_{\mu\nu}\; F_A^{\mu \nu} + \frac{1}{2} D^{ab}_\mu \phi_b \; D_{ab}^\mu \phi^b + i \psi_i^\dagger \sigma^\mu D^{ij}_\mu \psi_j  - ( m_{ij}\, \psi_i \,(i \sigma_2)\, \psi_j + \mathrm{h.c.} )\nonumber \\
&- ( Y^a_{ij} \phi_a \psi_i \,(i \sigma_2)\, \psi_j + \mathrm{h.c.} ) - J_a \;\phi^a - m_{ab}\; \phi^a \phi^b - h_{abc}\; \phi^a \phi^b \phi^c \nonumber\\ &- \lambda_{abcd}\; \phi^a \phi^b \phi^c \phi^d  + \mathcal{L}_{gf} + \mathcal{L}_{gh}
\end{align}
It shall be denoted that no artificial symmetrization on indices is assumed, as this is the case in \cite{LuoWangXiao,MV1,MV2,MV3,vev1,vev2,gauge3L,gauge3L2}.
Renormalization group equations may be generated up to two loop level for gauge couplings, Yukawas, quartic and cubic scalar interactions as well as masses and anomalous dimensions for fermions and scalars. For simple gauge groups, 3-loop Beta functions for the gauge coupling is available as well.
The theory consists of Weyl fermions and real scalars. Nevertheless, ARGES can handle complex scalars and converts them to real compontent fields internally.
$\mathcal{L}_{gf}$ and $\mathcal{L}_{gh}$ represent the gauge fixing and ghost part respectively, however, most RGEs are gauge independent anyway with exception of anomalous dimensions and vacuum expectation values. For those, $R_\xi$ gauge is implicitly assumed, yielding:
\begin{align}
\mathcal{L}_{gf} = -\frac{1}{2 \xi} A^I_\mu \,A^{I \mu} 
\end{align}
Contributions from the gauge sector are expressed in form of the quadratic Casimir, the Dynkin indices and $\Lambda$ invariants:
\begin{align}
C_2(R)\;\delta_{ab} &= T^A_{ac} \;T^A_{cb} \label{casimir}\\
S_2(R) \;\delta^{AB} &= T^A_{ab} \; T^B_{ba} \label{dynkin}\\
\Lambda_{abcd} &= T^A_{ac} \; T^A_{bd} \label{lambda}
\end{align}
Where $T^A_{ab}$ denominate generator of the representation of the gauge subgroup for scalar, fermionic and vector fields. Since scalars are decomposed to be real, their generators are imaginary and antisymmetric in the indices $a \leftrightarrow b$. Since vectors are considered to be real and in the adjoint representation, their generator are imaginary, totally antisymmetric and exhibit a Jacobi identity.
\newpage
\section{Defining a model}
\subsection{General Remarks}
As it is a guiding principle of ARGES to handle unspecified ranks and generations, contraction over indices are not performed by matrix multiplication techniques, but unspecified sums over contraction functions. This gives the user full control over the index contractions, but also burdens the kernel to resolve difficult systems of unspecified sums. As it is the case throughout the code, pattern matching is employed to enhance the performance of each computation. Particularly, delta symbols may be resolved mostly by pattern matching. To properly utilize this enhancement, it is recommended to the user to use the build-in function $\mathtt{KroneckerDelta}$ when specifying contraction functions. 
\subsection{Gauge sector and initial Setup}
The initial step when writing a model file is to invoke the function $\mathtt{Reset[\;];}$ to initialize background variables. This command also clears the package and allows a fresh and redefinition of a model. This is particularly useful when accidentally adding fields or couplings twice. \\
A second step is the to define the variable $\mathtt{NumberOfSubgroups}$, specifying the number of gauge subgroups. The default for this variable is set to 1, but it can be set to any non-negative integer, including zero. The value assigned to $\mathtt{NumberOfSubgroups}$ cannot be a symbol.
\vspace{1em}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
Reset[];
NumberOfSubgroups=1;
\end{lstlisting}
\vspace{1em}
In order to define gauge groups and their respective fields the function $\mathtt{Gauge[\dots]}$ must be invoded $\mathtt{NumberOfSubgroups}$ times. The syntax is:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
Gauge[Symbol, Algebra, Rank, {Charge/Representation1, ...}];
\end{lstlisting}
The last argument is a list of the charges (in case of $\mathrm{U}(1)$ subgroups) or representations of gauge bosons. The first charge/representation will be associated with the group first declared with $\mathtt{Gauge[\dots]}$, and so on. The dimension of this list must match $\mathtt{NumberOfSubgroups}$. ARGES is agnostic regarding anti-representations, only positive numbers are valid. However, charges, representations and rank of the subgroup may be declared as a variable. \newline Currently available algebras and ranks are $\mathrm{U}(1)$ , $\mathrm{SU}(N_U)$, and $\mathrm{SO}(N_O)$ with $N_U > 1$ and $N_O > 2$. Gauge fields are assumed to be in the adjoint representation of their own subgroup and singulets in the others. Kinetic mixing is not implemented. \newline Unknown Gauge groups are allowed, but their invariants can of course not be further resolved.
\FloatBarrier
\subsection{Fermionic fields}
Fermions may be defined via $\mathtt{WeylFermion}$:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
WeylFermion[Symbol, Generations, {Charge/Representation1, ...}];
\end{lstlisting}
The same comments as for $\mathtt{Gauge[\dots]}$ apply here for the last argument, including the fact that the representations and charges may be defined as variables. The same is true for the number of generations.
\FloatBarrier
\subsection{Scalar fields}
Scalars may be defined as real or complex, using the following declarations:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
RealScalar[Symbol, {Gen1, Gen2}, {Charge/Representation1, ...}];
ComplexScalar[Symbol, {Gen1, Gen2}, {Charge/Representation1, ...}];
\end{lstlisting}
While most of the syntax matches the fermionic case, scalars are natively matrix fields and posses two generation indices. For a vector-like generation structure, one of them may be set to 1.\newline
Internally, all complex scalars are expanded into real components, using the decomposition $\Phi = \frac{1}{\sqrt{2}} \left(\mathrm{Re} \, \Phi + i \, \mathrm{Im} \,\Phi\right)$. These component fields may be referenced as $\mathtt{Re[\Phi]}$ and $\mathtt{Im[\Phi]}$ when defining interactions.
\FloatBarrier
\subsection{Yukawa interactions}
Two possibilities exist to define Yukawa interactions. One is the classical approach as a matrix in the fermionic generation indices, the syntax looks like this:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
YukawaYaij[Symbol, Scalar, Ferm1, Ferm2, {Contraction1, ...}, Factor];
\end{lstlisting}
The arguments two to four are the symbols of the fields involved. The Scalar entered may be real (or the real/imaginary part of complex field - using $\mathtt{Re[...]}$/$\mathtt{Im[...]}$) or complex. The hermitian conjugate of a complex field may be used as well, utilizing $\mathtt{adj[...]}$. The latter applies both to the fermions and scalars. The last argument is a prefactor which might be complex, but its usage is optional, and the default value 1. The fifth arguments is to define the contraction between the gauge indices of the fields involved for this term. This list has $\mathtt{NumberOfSubgroups}$ elements and contains functions with three arguments, where which correspond to the gauge indices of the scalar, first and second fermion, strictly in this order. The hermitian conjugate of this Yukawa term is added automatically, no further prefactor or symmetrization will be employed. There is no contraction for scalar generations for these terms, but traces and products in the fermionic indices are denoted as $\mathtt{tr[...]}$ and $\mathtt{prod[...][i,j]}$ respectively.\newline 
Furthermore, Yukawa couplings may be defined as scalars:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
YukawaY[Sym, Scalar, Ferm1, Ferm2, {Contr1, ...}, GenerationContr];
\end{lstlisting}
Which is the same thing with exception of the now mandatory last argument, which is a now a function with four arguments, and defines the contraction between the first and second generation index of the scalar, and the generation indices of the first and second fermion, strictly in that order. Again, the hermitian conjugate is added without further symmetry factors. Any Yukawa coupling is generally considered complex.
\FloatBarrier
\subsection{Scalar quartic interactions}
Scalar quartics may be defined via:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
Quartic$\lambda$abcd[Sym, Sc1, Sc2, Sc3, Sc4, {Contr1, ...}, GenerationContr];
\end{lstlisting}
Where arguments two to five are symbols of the scalars involved. Again, real and complex scalars may be entered. It may be denoted that for complex fields, prefactors of the real composite fields are added automatically. The sixth element is a list of gauge index contractions, taking four arguments reflecting the four scalar fields. The generation contraction on the last position is a function of eight arguments, where the first two are for the two generation indices of the first scalar field and so on. \newline No implicit factor 4! is added, so each symmetry factor needs to be included in the last argument. However, for calculational purposes, the quartics are completely symmetrized under permutation of the scalar fields, and additional factors are added to preserve the norm. The final prefactor may be checked using the routine to output beta function, as it is explained in the next chapter.
\FloatBarrier
\subsection{Vacuum expectation values}
For theories developing VEVs of scalars, these have to be declared before using:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
VEV[Symbol, Scalar, {Gen1, Gen2}, {GaugeIndex1, ...}, Factor];
\end{lstlisting}
Where the first two arguments specify the symbols of the VEV and the scalar field. Since VEVs are considered to be real, real component fields need to be utilized for the scalars. However, its is of course possible to define imaginary components to develop a VEV. The third and fourth arguments are Lists containing the generation and gauge indices of the field in the usual fashion. The last argument is an additional factor for the VEV, however, factors already defined for the component fields are implicitly assumed for the VEV as well. This argument is optional and defaults to one. In general, to facilitate more complex vacuum structures, multiple VEVs with the same symbol, but other arguments or different VEVs with the same arguments may be specified, these are always assumed in an additive manner.
\FloatBarrier
\subsection{Scalar cubic interactions}
Cubic terms involving scalars may be added in a similar fashion using:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
CubicHabc[Symbol, Sc1, Sc2, Sc3, GaugeContractions, GenContraction];
\end{lstlisting}
Here again, complex or real scalars are accepted as input. Gauge contractions are functions of three arguments, which are the gauge indices of the involved fields, while the generation contraction accepts six arguments, two subsequently as the generation indices of one scalar. Internally, cubics and scalar masses are obtained by introducing one and two dummy fields into the quartic RGEs. Hence, those parameters inherit its property be symmetrized but not to have symmetry factors in the Lagrangian on its own. Fermion masses are obtained in a similar fashion from Yukawa interactions. Hence, the RGE subsystem of these superrenormalizable parameters decouples from the one of renormalizable parameters.
\subsection{Scalar Mass}
Scalar masses may be added using:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
ScalarMassMab[Symbol, Sc1, Sc2, GaugeContractions, GenContraction];
\end{lstlisting}
All remarks on scalar quartics and cubics apply in this context as well. Including the fact that there is no build-in symmetry factor when having a mass term diagonal in the scalar eigenstates. These need to be added manually.
\subsection{Fermion Mass}
Fermion masses are obtained from introducing a dummy scalar field into a Yukawa interaction, and hence share its symmetry factors. Furthermore, there are two versions as well, representing Fermion mass matrices and single masses:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
FermionMassMij[Symbol, Ferm1, Ferm2, {Contraction1, ...}, Factor];
FermionMassM[Symbol, Ferm1, Ferm2, {Contr1, ...}, GenerationContr];
\end{lstlisting}
Most comments from the Yukawa section still apply, however, gauge and generation contractions only expect two arguments, which are the Fermion indices.
\subsection{Scalar linear interactions}
For scalars with linear interactions at tree level, these may be added using a notation similar to scalar quartic, cubic and mass interactions:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
ScalarLinearJa[Symbol, Scalar, GaugeContractions, GenContraction];
\end{lstlisting}
Here, gauge contractions only take one argument, and Generation contraction two - which corresponds to the indices of the involved scalar.
\newpage
\section{Output}
\subsection{General remarks}
Beta functions are in general calculated on demand, the user has full control over which function to calculate and to which loop order. The general syntax follows $\mathtt{\beta[..., Loop Order]}$. Other inputs may be the involved fields and their quantum numbers. If the loop order is set to 0, the linear combination of couplings is returned as written in the Lagrangian modulo the fields and indices specified. This is particularly useful to find out prefactors after symmetrization and from taking real and imaginary parts of complex scalars. This is also the main mechanism to disentangle the running of couplings.\newline The gauge and generation indices of the involved fields are entered as lists. For abelian gauge groups, it is necessary to insert 1 instead of the charge. If the representation, generations or rank are variables (for both the definition or the input indices), the situation is slightly more complicated, since the Mathematica kernel tries to cover all special cases when handling sums with variables, and drags them through the entire computation. This may slow the calculation severely or even break it. ARGES utilizes the global variable \$\texttt{Assumptions} (special care must be taken when using this outside of ARGES, since \texttt{Reset[]} wipes it completely) to refine these cases. If variables appear in the definition, the index 1 is always safe as input, for higher numbers, a global assumption must be appended to respecify the range of the variables. If there is an input of variables, their range as well as the fact that these are integers must be appended as well. \newline 
Gauge invariants are by default not resolved. To do so, the function $\mathtt{ComputeInvariants[\;]}$ must first be evoked, which fills the replacement rule list \texttt{subInvariants} with appropriate substitutions. It shall be denoted that for scalars, at two loop in the Yukawa betas and one loop in the quartic betas, as well as for fermions at two loop in the quartic betas, gauge terms appear which may not be expressed by Casimirs and Dynkin Indices, but rather: $ \Lambda_{ijkl} = T ^a_{ik}\;T^a_{jl}$, where $T^a_{ij}$ are generators of a gauge subgroup for one particle. However, closed formulas for any rank may not exist for all representations, as it is the case for instance for the fundamental representation of SU(N).
\subsection{Gauge coupling beta function}
For gauge couplings, the beta functions may be calculated via:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Gauge, LoopOrder]
$\beta$[$\alpha$[Gauge], LoopOrder]
\end{lstlisting}
Where the second version returns the beta function of and with $ \alpha(g) = \frac{g^2}{(4 \pi)^2}$. In any case, the functions still contain Casimirs and Dynkin Indices. To resolve these (when the Algebra and the representation is specified), the substitution \texttt{subInvariants} may be utilized.
\FloatBarrier\newpage
\subsection{Yukawa beta function}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Scalar, Ferm1, Ferm2, SIdxList, F1IdxList, F2IdxList, LoopOrder]
\end{lstlisting}
The first three arguments are symbols of the scalar and the fermions involved, the fourth argument is a list of $\mathtt{NumberOfSubgroups} + 2$ elements, where the first two mark the generations of the scalar fields, and the rest its gauge indices. In the same manner, the fifth and sixth arguments are lists of $\mathtt{NumberOfSubgroups} + 1$ elements, where the first argument is the generation, and the rest gauge indices for each fermion respectively.\newline\FloatBarrier 
\subsection{Scalar quartic beta function}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Sc1, Sc2, Sc3, Sc4, S1List, S2List, S3List, S4List, LoopOrder]
\end{lstlisting}
Following the notation before, the first four symbols mark the scalar fields, while the arguments five to eight are their respective indices, packed in lists of $\mathtt{NumberOfSubgroups} + 2$ elements, with the first two being the generation and the rest the gauge indices. \newline \FloatBarrier
\subsection{Vacuum expectation value beta function}
The running of vacuum expectation values is obtained via:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[vev, LoopOrder]
\end{lstlisting}
It may be denoted that these are in general gauge depended. Here, all gauge groups are assumed have the same constant $\xi$ for $R_{\xi}$ gauge fixing.\FloatBarrier
\subsection{Scalar cubic beta functions}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Sc1, Sc2, Sc3, S1List, S2List, S3List, LoopOrder]
\end{lstlisting}
The input is similar to the one for the quartic beta function. Fermion masses also contribute to this RGE from one-loop level on, however, scalar masses do not.\\ \FloatBarrier
\subsection{Scalar mass beta functions}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Sc1, Sc2, S1List, S2List, LoopOrder]
\end{lstlisting}
The input is similar to the one for the sclar quartic or cubic beta functions. Fermion masses and scalar cubics contribute to this RGE from one-loop level on.\\ \FloatBarrier
\subsection{Scalar linear interaction beta functions}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Scalar, ScalarIndexList, LoopOrder]
\end{lstlisting}
Here again, the input follows the previous beta functions. From one loop level on, cubic interactions, fermion and scalar masses may appear. Even here, the RGE should be checked for possible prefactors due to symmetrization, using $\mathtt{LoopOrder}$ zero.  \\ \FloatBarrier
\subsection{Fermion mass beta function}
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\beta$[Ferm1, Ferm2, F1IndexList, F2IndexList, LoopOrder]
\end{lstlisting}
The notation follows the one for Yukawa beta functions quite closely. Scalar quartics contribute to these RGEs form two-loop level on, but it is independent from scalar masses. \\
\subsection{Fermion anomalous dimension}
Anomalous dimensions for fermionic fields may be computed via:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\gamma$[Ferm1, Ferm2, F1IndexList, F2IndexList, LoopOrder]
\end{lstlisting}
Here, the arguments are the same as for the fermionic mass. If $\mathtt{LoopLevel}$ is set to zero, a product of delta symbols over the particle type, generation and gauge indices is returned. The result is in general gauge depended and assumed in a the $R_\xi$ gauge for all couplings. \\ \FloatBarrier
\subsection{Scalar anomalous dimension}
Anomalous dimensions for scalar fields are very similar to the fermionic case:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
$\gamma$[Scalar1, Scalar2, S1IndexList, S2IndexList, LoopOrder]
\end{lstlisting}
Only that scalars (which are again automatically decomposed) are added for the first two arguments, and the second two contain $\mathtt{NumberOfGaugegroups+2}$ elements due to the additional generation indices. If $\mathtt{LoopLevel}$ is set to zero, a product of delta symbols over the particle type, generation and gauge indices is returned. The result is in general gauge depended and assumed in a the $R_\xi$ gauge for all couplings. \\ \FloatBarrier
\newpage
\section{Advanced Options}
\subsection{Advanced Simplification}
By default, ARGES resolves index contraction by explicitly summing over these. However, especially when dealing with indefinite bounds or non-numerical external indices, the Mathematica kernel runs into trouble resolving these sums, or fills up the memory be dragging a plethora of conditions with the solutions. One possibility to overcome is to append conditions for non-numerical bounds or indices to Mathematica standard assumption condition $\$\mathtt{Assumptions}$. \\ In favour of wasting system resources by performing nested sums, advanced pattern matching techniques are employed by ARGES to simplify any index sum before residing to actual summation. In fact, two versions of this algorithm exist. One to employ with every index sum calculated, which is deliberately kept lightweight and inaccessible to the user to protect the performance of ARGES. The second one is automatically performed at the end of each calculation to simplify the result. It is accessible by the user through:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
SimplifyProduct[expression]
\end{lstlisting}
The following simplifications are implemented in both versions of the algorithms:
\begin{itemize}
\item the expression is expanded and factor dragged under the sum
\item trivial summations $\sum_{i=1}^1$ are resolved
\item summations like $\sum_{i=1}^{k} \mathtt{KroneckerDelta}[i,j]$ are resolved - here it is implicitly assumed that $j \leq k$; it is hence wise to formulate contraction rules using $\mathtt{KroneckerDelta}[...]$
\item some simplifications for terms containing the $\mathtt{Generator}$ keyword are applied - refer to the next section for more details
\item simplify consecutive operations of $\mathtt{conj}$, $\mathtt{transpose}$, $\mathtt{adj}$
\item reorder and simplify Yukawa traces 
\end{itemize}
While the version accessible to $\mathtt{SimplifyProduct}$ additionally may:
\begin{itemize}
\item simplify expressions with the $\mathtt{Generator}$ keyword much further
\item simplify $\Lambda_{abbc} = C_2 \,\delta_{ac}$ for fermions
\item employ user specified simplification rules
\end{itemize}
Indeed, one may add additional assumption to the simplification rules for $\mathtt{SimplifyProduct}$ by defining the variable $\mathtt{subSimplifySum}$. Any rules defined within those should assume that every $\mathtt{Sum}$ keyword is rebranded as $\mathtt{SimplifySum}$, an example could look like this:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
subSimplifySum = {
   ...
   SimplifySum[fac_ func[x_], Sum1___, {x_, 1, xmax_}, Sum2___] :> 
      SimplifySum[ fac simplefunc, Sum1, Sum2],
   ...
}
\end{lstlisting}
\subsection{The $\mathtt{Generator}$ keyword}
In order to handle adjoint representations better, ARGES allows to use the following structure
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
Generator[Args][a, i, j]
\end{lstlisting}
Which represents a generator of the gauge group $T^a_{ij}$ (note that here, indices do not run over other gauge groups, generations or particle types), where $\mathtt{Args}$ marks arbitrary argument which can be chosen by the user, however, the canonical way in ARGES is to choose $\mathtt{Generator}[\mathtt{FieldSymbol},\;\mathtt{GaugeSymbol}]$ or $\mathtt{Generator}[\mathtt{GaugeSymbol}]$ for the adjoint representation.
The following relations are build into the simplification algorithm in ARGES:
\begin{itemize}
\item ${T^a_{ij}}^* = T^a_{ji}$
\item $T^a_{ij} T^a_{jk} = C_2\, \delta_{ik}$
\item $T^a_{ij} T^b_{ji} = T_2\, \delta^{ab}$
\end{itemize}
Where $T_2$ is easily translated into $S_2$ employing $\mathtt{subInvariants}$. Furthermore, for $\mathtt{SimplifyProduct}$, the relation
\begin{align*}
T^a_{ij} T^b_{jk} T^a_{kl} = \left(C_2(R) - \dfrac{1}{2} C_2(G)\right) T^b_{il}
\end{align*}
is also included.
\newpage
\section{Example: SM with $N_c$ colors and third generation fermions}
In this section, we will present an example of usage featuring the Standard Model with $N_c$ colors and only considering third generation fermions. The model file looks like this (depicted in plain text without notebook specific replacements):
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
(* SM with Nc colors and 3rd generation fermions only *)
Needs["ARGES`"]

(* Reset ARGES every time reading in this file *)
Reset[];

(* Declare number of gauge groups first *)
NumberOfSubgroups = 3;

(* Specify gauge groups, no GUT normalization *)
Gauge[gy, U, 1, {0, 1, 1}];
Gauge[g, SU, 2, {0, 3, 1}];
Gauge[gs, SU, Nc, {0, 1, Sqr[Nc] - 1}];

(* Specify full fermion sector *)
(* we will define no Yukawas for the first 2 generations later *)
WeylFermion[Q, 3, {+1/6, 2, Nc}];
WeylFermion[L, 3, {-1/2, 2, 1}];
WeylFermion[D, 3, {-1/3, 1, Nc}];
WeylFermion[U, 3, {+2/3, 1, Nc}];
WeylFermion[E, 3, {-1, 1, 1}];

(* Scalar Field with VEV *)
ComplexScalar[H, {1, 1}, {+1/2, 2, 1}];
VEV[v, Re[H], {1, 1}, {1, 2, 1}, 1];

(* Yukawa sector - third generations only *)
YukawaY[ yb, H, adj[Q], D, 
 {1 &, KroneckerDelta[#1, #2] &, KroneckerDelta[#2, #3] &}, 
 (KroneckerDelta[#3, 3] KroneckerDelta[#4, 3])& ];
 
YukawaY[ ytau, H, adj[L], E, 
 {1 &, KroneckerDelta[#1, #2] &, 1 &}, 
 (KroneckerDelta[#3, 3] KroneckerDelta[#4, 3])& ];
 
YukawaY[ yt, adj[H], adj[Q], U, 
 {1 &, Eps[#1, #2] &, KroneckerDelta[#2, #3] &},
 (KroneckerDelta[#3, 3] KroneckerDelta[#4, 3])& ];

(* Quartic coupling ~ -1/2 lamda |phi|^4 *)
Quartic\[Lambda]abcd[\[Lambda], adj[H], H, adj[H], H, 
 {1&, (KroneckerDelta[#1, #2] KroneckerDelta[#3, #4])&, 1&}, 
 (KroneckerDelta[#2, #3] KroneckerDelta[#1, #4] 
  KroneckerDelta[#5, #8] KroneckerDelta[#6, #7])/2 & ];

(* Compute gauge invariants already *)
ComputeInvariants[];
\end{lstlisting}
This file is included in the bundle as $\mathtt{ARGES/models/SM\_Nc\_3rd.m}$. It shall be denoted that $\mathtt{Eps}$ and $\mathtt{Sqr}$ are keywords defined by ARGES for convenience, while $\mathtt{KroneckerDelta}$ is part of the Wolfram Language. It is less cumbersome to utilize the support for anonymous ('pure') functions as done in the example.
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
(* non-anonymous definition *)
f[x1_, x2_, ...] := g[x1, x2, ...];

(* anonymous definition using pure functions *)
f := g[#1, #2, ...] & ;
f := Function[{x1, x2, ...}, g[x1, x2, ...]]; 
\end{lstlisting}
Now the model can be called from a batch file or interactively from a notebook:
\begin{lstlisting}[language=mathematica,mathescape,columns=flexible,backgroundcolor=\color{light-gray}]
In[1]:= Get["SM_Nc_3rd.m", Path -> {"~/ARGES/models/"}]
        
        (* real matrix elements, substitute gauge invariants *)        
In[2]:= subs = Join[subInvariants, 
        {conj[yt] -> yt, conj[yb] -> yb, conj[ytau] -> ytau}];

        (* gauge couplings beta functions - here just gs for brevity *)        
In[3]:= Power[4 \[Pi], 2] \[Beta][gs, 1] //. subs // Expand
Out[3]= 4 gs^3 - (11 gs^3 Nc)/3

In[4]:= Power[4 \[Pi], 4] \[Beta][gs, 2] //. subs // Expand
Out[4]= (9 g^2 gs^3)/2 + (11 gs^3 gy^2)/6 - (6 gs^5)/Nc + 26 gs^5 Nc - 
        (34 gs^5 Nc^2)/3 + (2 gs^3 yb^2)/(-1 + Nc^2) - 
        (2 gs^3 Nc^2 yb^2)/(-1 + Nc^2) + (2 gs^3 yt^2)/(-1 + Nc^2) - 
        (2 gs^3 Nc^2 yt^2)/(-1 + Nc^2)
        
        (* Yukawa yt coupling: get symmetry factors first *)
In[5]:=  \[Beta][H, adj[Q], U, {1, 1, 1, 1, 1}, {3, 1, 2, 1}, 
         {3, 1, 1, 1}, 0] //. subs // Expand
Out[5]= yt

In[6]:= Power[4 \[Pi], 2] \[Beta][H, adj[Q], U, {1, 1, 1, 1, 1}, 
        {3, 1, 2, 1}, {3, 1, 1, 1}, 1] //. subs // Expand
Out[6]= -((9 g^2 yt)/4) - (17 gy^2 yt)/12 + (3 gs^2 yt)/Nc - 
        3 gs^2 Nc yt - (3 yb^2 yt)/2 + Nc yb^2 yt + (3 yt^3)/2 + 
        Nc yt^3 + yt ytau^2
        
In[7]:= Power[4 \[Pi], 4] \[Beta][H, adj[Q], U, {1, 1, 1, 1, 1}, 
        {3, 1, 2, 1}, {3, 1, 1, 1}, 2] //. subs // Expand
Out[7]= -8 g^4 yt + (53 gs^4 yt)/3 - 3/4 g^2 gy^2 yt + (61 gy^4 yt)/24 - 
        (3 gs^4 yt)/(4 Nc^2) - (27 g^2 gs^2 yt)/(8 Nc) - (10 gs^4 yt)/Nc
        - (19 gs^2 gy^2 yt)/(24 Nc) + 3/4 g^4 Nc yt + 27/8 g^2 gs^2 Nc yt
        + 10 gs^4 Nc yt + 19/24 gs^2 gy^2 Nc yt + 319/324 gy^4 Nc yt - 
        203/12 gs^4 Nc^2 yt + 9/16 g^2 yb^2 yt - 5/2 gs^2 yb^2 yt - 
        43/48 gy^2 yb^2 yt + (6 gs^2 yb^2 yt)/Nc + 15/8 g^2 Nc yb^2 yt -
        6 gs^2 Nc yb^2 yt + 25/72 gy^2 Nc yb^2 yt + 5/2 gs^2 Nc^2 yb^2 yt
        + (11 yb^4 yt)/4 - Nc yb^4 yt + (135 g^2 yt^3)/16 - 
        (5 gs^2 yt^3)/2 + (223 gy^2 yt^3)/48 - (6 gs^2 yt^3)/Nc + 
        15/8 g^2 Nc yt^3 + 6 gs^2 Nc yt^3 + 85/72 gy^2 Nc yt^3 + 
        5/2 gs^2 Nc^2 yt^3 - (5 yb^2 yt^3)/4 - 1/2 Nc yb^2 yt^3 + 
        (3 yt^5)/2 - (9 Nc yt^5)/2 + 15/8 g^2 yt ytau^2 + 
        25/8 gy^2 yt ytau^2 + 5/4 yb^2 yt ytau^2 - (9 yt^3 ytau^2)/4 - 
        (9 yt ytau^4)/4 - 6 yt^3 \[Lambda] + (3 yt \[Lambda]^2)/2
        
        (* VEV *)
In[8]:= Power[4 \[Pi], 2] \[Beta][v, 1] //. subs // Expand
Out[8]= (9 g^2 v)/4 + (3 gy^2 v)/4 - Nc v yb^2 - Nc v yt^2 - v ytau^2 + 
        3/4 g^2 v \[Xi] + 1/4 gy^2 v \[Xi]
        
In[9]:= Power[4 \[Pi], 4] \[Beta][v, 2] //. subs // Expand
Out[9]= (451 g^4 v)/32 - 9/16 g^2 gy^2 v - (211 gy^4 v)/96 - 
        15/8 g^4 Nc v - 55/72 gy^4 Nc v + 5/2 gs^2 v yb^2 - 
        15/8 g^2 Nc v yb^2 - 25/72 gy^2 Nc v yb^2 - 
        5/2 gs^2 Nc^2 v yb^2 + 9/4 Nc v yb^4 + 5/2 gs^2 v yt^2 - 
        15/8 g^2 Nc v yt^2 - 85/72 gy^2 Nc v yt^2 - 5/2 gs^2 Nc^2 v yt^2 
        - 1/2 Nc v yb^2 yt^2 + 9/4 Nc v yt^4 - 15/8 g^2 v ytau^2 - 
        25/8 gy^2 v ytau^2 + (9 v ytau^4)/4 - (3 v \[Lambda]^2)/2 + 
        27/8 g^4 v \[Xi] + 3/4 g^2 gy^2 v \[Xi] + 1/8 gy^4 v \[Xi] - 
        3/2 g^2 Nc v yb^2 \[Xi] - 1/2 gy^2 Nc v yb^2 \[Xi] - 
        3/2 g^2 Nc v yt^2 \[Xi] - 1/2 gy^2 Nc v yt^2 \[Xi] - 
        3/2 g^2 v ytau^2 \[Xi] - 1/2 gy^2 v ytau^2 \[Xi] + 
        3/4 g^2 gy^2 v \[Xi]^2 + 1/8 gy^4 v \[Xi]^2

         (* Quartic *)
In[10]:= \[Beta][adj[H], H, adj[H], H, {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, 
         {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, 0]
Out[10]= \[Lambda]/2

In[11]:= 2 Power[4 \[Pi], 2] \[Beta][adj[H], H, adj[H], H, 
         {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, 
         {1, 1, 1, 1, 1}, 1] //. subs // Expand
Out[11]= (9 g^4)/4 + (3 g^2 gy^2)/2 + (3 gy^4)/4 - 4 Nc yb^4 - 4 Nc yt^4 
         - 4 ytau^4 - 9 g^2 \[Lambda] - 3 gy^2 \[Lambda] + 
         4 Nc yb^2 \[Lambda] + 4 Nc yt^2 \[Lambda] + 4 ytau^2 \[Lambda] 
         + 12 \[Lambda]^2

In[12]:= 2 Power[4 \[Pi], 4] \[Beta][adj[H], H, adj[H], H, 
         {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, {1, 1, 1, 1, 1}, 
         {1, 1, 1, 1, 1}, 2] //. subs // Expand         
Out[12]= (449 g^6)/8 - (145 g^4 gy^2)/24 - (383 g^2 gy^4)/24 - 
         (203 gy^6)/24 - 6 g^6 Nc - 2 g^4 gy^2 Nc - 22/9 g^2 gy^4 Nc - 
         (22 gy^6 Nc)/9 - 3/2 g^4 Nc yb^2 + 3 g^2 gy^2 Nc yb^2 + 
         5/6 gy^4 Nc yb^2 + 8 gs^2 yb^4 + 8/9 gy^2 Nc yb^4 - 
         8 gs^2 Nc^2 yb^4 + 20 Nc yb^6 - 3/2 g^4 Nc yt^2 + 
         7 g^2 gy^2 Nc yt^2 - 19/6 gy^4 Nc yt^2 - 4 Nc yb^4 yt^2 + 
         8 gs^2 yt^4 - 16/9 gy^2 Nc yt^4 - 8 gs^2 Nc^2 yt^4 - 
         4 Nc yb^2 yt^4 + 20 Nc yt^6 - (3 g^4 ytau^2)/2 + 
         11 g^2 gy^2 ytau^2 - (25 gy^4 ytau^2)/2 - 8 gy^2 ytau^4 + 
         20 ytau^6 - (253 g^4 \[Lambda])/8 + 39/4 g^2 gy^2 \[Lambda] + 
         (409 gy^4 \[Lambda])/24 + 15/2 g^4 Nc \[Lambda] + 
         55/18 gy^4 Nc \[Lambda] - 10 gs^2 yb^2 \[Lambda] + 
         15/2 g^2 Nc yb^2 \[Lambda] + 25/18 gy^2 Nc yb^2 \[Lambda] + 
         10 gs^2 Nc^2 yb^2 \[Lambda] - Nc yb^4 \[Lambda] - 
         10 gs^2 yt^2 \[Lambda] + 15/2 g^2 Nc yt^2 \[Lambda] + 
         85/18 gy^2 Nc yt^2 \[Lambda] + 10 gs^2 Nc^2 yt^2 \[Lambda] - 
         14 Nc yb^2 yt^2 \[Lambda] - Nc yt^4 \[Lambda] + 
         15/2 g^2 ytau^2 \[Lambda] + 25/2 gy^2 ytau^2 \[Lambda] - 
         ytau^4 \[Lambda] + 54 g^2 \[Lambda]^2 + 18 gy^2 \[Lambda]^2 - 
         24 Nc yb^2 \[Lambda]^2 - 24 Nc yt^2 \[Lambda]^2 - 
         24 ytau^2 \[Lambda]^2 - 78 \[Lambda]^3
         
\end{lstlisting}
\newpage
\printbibliography
\end{document}